NumPy---Array Programming in Python

### 2020-02-04 body.tex

\emph{Strides} are necessary to interpret computer memory, which stores elements
linearly, as multidimensional arrays.
It describes the number of bytes to move forward in memory to jump from row to
row or column to column.
Consider, for example, a 2-D array of floating-point numbers with shape
$(4, 3)$, where each element occupies 8 bytes in memory.
To move between consecutive columns, we need to jump forward 8 bytes in memory,
and to access the next row $3 \times 8 = 24$ bytes.
The strides of that array are therefore $(24, 8)$.  NumPy can
store arrays in either C or Fortran memory order, iterating
first over either rows or columns.  This allows external libraries
written in those languages to access NumPy array data in memory directly.

@@

The interactive environment created by this array programming
foundation---along with the surrounding ecosystem of tools---inside of
IPython or Jupyter, is ideally suited to exploratory data analysis.
Users fluidly inspect, manipulate, and visualize their data, and
rapidly iterate to refine programming statements. These statements are
then stitched together into imperative or functional programs or
notebooks containing both computation and narrative.
Scientific computing beyond exploratory work is often done in a text editor
or an integrated development environment (IDEs) such as Spyder\cite{spyder2020}.
This rich and productive environment has made Python popular
for scientific research.

@@

NumPy provides in-memory, multidimensional, homogeneously typed
arrays on CPUs.  It runs on machines
ranging from embedded devices to the world's largest supercomputers,
with performance approaching that of compiled languages.  For the
most its existence, NumPy addressed the vast majority of
array computation use cases.

@@

In recent years, there has been a rapid growth in the development of
specialized hardware, including graphics processing
units (GPUs), tensor processing units (TPUs), and field-programmable
gate arrays (FPGAs).  Data sets routinely exceed the memory capacity of a single machine and may be stored on multiple machines or in
the cloud.  Due to its in-memory data model, NumPy is currently unable to
utilize such specialized hardware and storage directly.  However, both
distributed data and the parallel execution of GPUs, TPUs, and FPGAs map well
to the \emph{paradigm} of array programming: a gap, therefore, existed between
available modern hardware architectures and the tools necessary to
leverage their computational power.

@@

Such libraries often mimic the NumPy API, because it lowers the
barrier to entry for newcomers and provides the wider community with a
stable array programming interface. This, in turn, prevents disruptive
schisms like the divergence of Numeric and NumArray.
But exploring new ways of working with arrays is experimental by nature
and, in fact, several promising libraries---such as Theano and Caffe---have
already ceased development. And each time that a user
decides to try new technology, they must
change import statements and ensure that the new library implements
all the parts of the NumPy API they currently use.

@@

Ideally, operating on specialized arrays using NumPy functions or semantics would
simply work, so that users could write code once, and would then benefit
from switching between NumPy arrays, GPU arrays, or distributed arrays, as appropriate.
To support array operations between external array objects, NumPy,
therefore, added the capability to act as a central coordination
mechanism with a well-specified API (Fig.~\ref{fig:ecosystem}).

@@

NumPy has subsequently seen prolific adoption, with millions of users,
first from research and later also industry.
It spurred the development of the broader scientific Python ecosystem and
heralded the current era of widespread use of Python for scientific computing.
The reasons for its popularity are varied and complex but include the fact that
NumPy (a) is built on top of Python,
(b) has a culture of collaboration and good software development practices, and
(c) provides a cutting-edge array programming environment for scientific data analysis.

>>>>

\emph{Strides} are necessary to interpret computer memory, which stores elements
linearly, as multidimensional arrays.
It describes the number of bytes to move forward in memory to jump from row to
row, column to column, and so forth.
Consider, for example, a 2-D array of floating-point numbers with shape
$(4, 3)$, where each element occupies 8 bytes in memory.
To move between consecutive columns, we need to jump forward 8 bytes in memory,
and to access the next row $3 \times 8 = 24$ bytes.
The strides of that array are therefore $(24, 8)$.  NumPy can
store arrays in either C or Fortran memory order, iterating
first over either rows or columns.  This allows external libraries
written in those languages to access NumPy array data in memory directly.

@@

The interactive environment created by this array programming
foundation---along with the surrounding ecosystem of tools---inside of
IPython or Jupyter, is ideally suited to exploratory data analysis.
Users fluidly inspect, manipulate, and visualize their data, and
rapidly iterate to refine programming statements. These statements are
then stitched together into imperative or functional programs, or
notebooks containing both computation and narrative.
Scientific computing beyond exploratory work is often done in a text editor
or an integrated development environment (IDEs) such as Spyder\cite{spyder2020}.
This rich and productive environment has made Python popular
for scientific research.

@@

NumPy provides in-memory, multidimensional, homogeneously typed
arrays on CPUs.  It runs on machines
ranging from embedded devices to the world's largest supercomputers,
with performance approaching that of compiled languages.  For
most its existence, NumPy addressed the vast majority of
array computation use cases.

@@

In recent years, there has been a rapid growth in the development of
specialized hardware, including graphics processing
units (GPUs), tensor processing units (TPUs), and field-programmable
gate arrays (FPGAs).  Data sets now routinely exceed the memory capacity of a single machine and may be stored on multiple machines or in
the cloud.  Due to its in-memory data model, NumPy is currently unable to
utilize such specialized hardware and storage directly.  However, both
distributed data and the parallel execution of GPUs, TPUs, and FPGAs map well
to the \emph{paradigm} of array programming: a gap, therefore, existed between
available modern hardware architectures and the tools necessary to
leverage their computational power.

@@

Such libraries often mimic the NumPy API, because it lowers the
barrier to entry for newcomers and provides the wider community with a
stable array programming interface. This, in turn, prevents disruptive
schisms like the divergence of Numeric and NumArray.
But exploring new ways of working with arrays is experimental by nature
and, in fact, several promising libraries---such as Theano and Caffe---have
already ceased development. And each time that a user
decides to try a new technology, they must
change import statements and ensure that the new library implements
all the parts of the NumPy API they currently use.

@@

Ideally, operating on specialized arrays using NumPy functions or semantics would
simply work, so that users could write code once, and would then benefit
from switching between NumPy arrays, GPU arrays, distributed arrays,
and so forth, as appropriate.
To support array operations between external array objects, NumPy,
therefore, added the capability to act as a central coordination
mechanism with a well-specified API (Fig.~\ref{fig:ecosystem}).

@@

NumPy has subsequently seen prolific adoption, with millions of users,
first from research and later also industry.
It spurred the development of the broader scientific Python ecosystem and
heralded the current era of widespread use of Python for scientific computing.
The reasons for its popularity are varied and complex but include the facts that
NumPy (a) is built on top of Python,
(b) has a culture of collaboration and good software development practices, and
(c) provides a cutting-edge array programming environment for scientific data analysis.
